### 2018-08-27 paper/finitecomplete-failures.tex
updates

For a given reference process $P$ and for each integer $p\ge 0$, we define a
CSP test process for failures refinement as shown below.
%
\begin{eqnarray}
U_F(p) & = & U_F(p,\varepsilon)
\\
U_F(p,s) & = & \big( \Extchoice e:(\Sigma - [P/s]^0) @ e \then \efail\then \Stop \big)
\label{eq:ufa}
\\ & & \extchoice \nonumber
\\ & & ([P/s]^0 = \varnothing)    \&   \big( \epass \then \Stop \big)
\label{eq:ufb}
\\ & & \extchoice \nonumber
\\ & & (\#s < p) \& \big( \Extchoice e:[P/s]^0 @ e \then U_F(p,s.e) \big)
\label{eq:ufc}
\\ & & \extchoice \nonumber
\\ & & (\#s = p) \& \big( \sqcap_{H\in\text{minHit}(P/s)} ( \Extchoice e:H @ e \then \epass \then\Stop   )  \big)
\label{eq:ufd}
\end{eqnarray}
%
A test is performed by running $U_F(p)$ concurrently with any SUT process $Q$
%\fixme{alcc: add an example?}\fxnote{jp: I think that the following explanations, together with examples in Section~\ref{sec:case}, suffice.} that operates on the same alphabet as $P$,
synchronising over alphabet $\Sigma$. Therefore, a test execution is any
trace of the concurrent process
\[
Q\parallel[\Sigma] U_F(p).
\]
It is assumed that the events $\efail$ and $\epass$, denoting FAIL and PASS
of the test execution, are events outside $\Sigma$. Since we assume that $Q$
is free of livelocks, it is guaranteed that each test execution terminates
after some $s\in\trc(P)$ with length $(p+1)$ at the latest. The test is
\emph{passed} by the SUT (written $Q\ \pass\ U_F(p)$) if, and only if, {\it
every} execution of $Q\parallel[\Sigma] U_F(p)$ terminates with PASS event
$\epass$. This can also be  expressed by means of a failures refinement.
\[
Q\ \pass\ U_F(p) \defs (\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(p)) \hide \Sigma
\]
This type of pass relation is often called \emph{must test}, because every
test execution must end with the $\epass$
event~\cite{Hennessy:1988:ATP:50497}. Note that it is necessary to use the
failures-refinement relation in this condition, and not the trace-refinement
relation:~$(Q\parallel[\Sigma] U_F(p)) \hide \Sigma$ may have  the same
visible traces $\varepsilon$ and $\langle \epass\rangle$ as the ``Test Passed
Process'' $(\epass\then\Stop)$. However, the former may nondeterministically
refuse $\epass$, due to a deadlock occurring when a faulty SUT process
executes concurrently with $U_F(p,s)$ executing branch (\ref{eq:ufd}),
because $\#s = p$. This is explained further in the next paragraphs.

>>>>

In the domain of process algebras, test cases are typically represented 
by processes interacting concurrently with the SUT 
process~\cite{Hennessy:1988:ATP:50497}. They synchronise with the SUT over its visible events and use some additional events outside the SUT's alphabet to express whether the test execution  passed or failed. %%%, or if no verdict could be obtained.

For a given reference process $P$ and for each integer $p\ge 0$, we define a
CSP test process for failures refinement as shown below.
%
\begin{eqnarray}
U_F(p) & = & U_F(p,\varepsilon)
\\
U_F(p,s) & = & \big( \Extchoice e:(\Sigma - [P/s]^0) @ e \then \efail\then \Stop \big)
\label{eq:ufa}
\\ & & \extchoice \nonumber
\\ & & ([P/s]^0 = \varnothing)    \&   \big( \epass \then \Stop \big)
\label{eq:ufb}
\\ & & \extchoice \nonumber
\\ & & (\#s < p) \& \big( \Extchoice e:[P/s]^0 @ e \then U_F(p,s.e) \big)
\label{eq:ufc}
\\ & & \extchoice \nonumber
\\ & & (\#s = p) \& \big( \sqcap_{H\in\text{minHit}(P/s)} ( \Extchoice e:H @ e \then \epass \then\Stop   )  \big)
\label{eq:ufd}
\end{eqnarray}
%
A test is performed by running $U_F(p)$ concurrently with any SUT process $Q$,
synchronising over alphabet $\Sigma$. Therefore, a test execution is any
trace of the concurrent process
\[
Q\parallel[\Sigma] U_F(p).
\]
It is assumed that the events $\efail$ and $\epass$, denoting FAIL and PASS
of the test execution, are events outside $\Sigma$. Since we assume that $Q$
is free of livelocks, it is guaranteed that each test execution terminates
after some $s\in\trc(P)$ with length $(p+1)$ at the latest. The test is
\emph{passed} by the SUT (written $Q\ \pass\ U_F(p)$) if, and only if, {\it
every} execution of $Q\parallel[\Sigma] U_F(p)$ terminates with PASS event
$\epass$. This can also be  expressed by means of a failures refinement.
\[
Q\ \pass\ U_F(p) \defs (\epass\then\Stop) \lessdet_F (Q\parallel[\Sigma] U_F(p)) \hide \Sigma
\]
This type of pass relation is often called \emph{must test}, because every
test execution must end with the $\epass$
event~\cite{Hennessy:1988:ATP:50497}. Note that it is necessary to use the
failures-refinement relation in this condition, and not the trace-refinement
relation:~$(Q\parallel[\Sigma] U_F(p)) \hide \Sigma$ may have  the same
visible traces $\varepsilon$ and $\langle \epass\rangle$ as the ``Test Passed
Process'' $(\epass\then\Stop)$. However, the former may nondeterministically
refuse $\epass$, due to a deadlock occurring when a faulty SUT process
executes concurrently with $U_F(p,s)$ executing branch (\ref{eq:ufd}),
because $\#s = p$. This is explained further in the next paragraphs.

Observe that the number of possible executions of $Q\parallel[\Sigma] U_F(p)$ is finite,
because the number of traces $s$ with maximal length $(p+1)$ is finite and the 
sets $[P/s]^0$, $(\Sigma - [P/s]^0)$, and $\text{minHit}(P/s)$ are finite.

### 2018-08-27 paper/introduction.tex
updates

While MBT is applied in different flavours, we consider the most effective
variant to be the one where test cases and concrete test data, as well as
checkers for the expected results~(\emph{test oracles}), are automatically
generated from a reference model. This guarantees maximal return of the
investment of time and effort to create the test model. The test suites
generated in this way, however, usually have different test strength,
depending on the generation algorithms applied.

For the safety-critical domain, test suites with guaranteed fault coverage
are of particular interest. For black-box testing, guarantees can be given
only if certain hypotheses are satisfied. These hypotheses are usually
specified by a \emph{fault domain}:~a set of models that may or may not
conform to the SUT. The so-called \emph{complete} test strategies guarantee
to uncover every conformance violation of the SUT with respect to a reference
model, provided that the true SUT behaviour is captured by a member of the
fault domain.

Generation methods for complete test suites have been developed for various
modelling formalisms. In this paper, we use \emph{Communicating Sequential
Processes (CSP)}~\cite{Hoare:1985:CSP:3921,Roscoe2010}. This is a mature
process-algebraic approach that has been shown to be well-suited for the
description of reactive control systems in many publications over almost five
decades. Industrial success has also been reported.

@@

This paper presents complete black-box test suites for software and systems
modelled using the {\it Communicating Sequential Processes (CSP)} process
algebra~\cite{Hoare:1985:CSP:3921,Roscoe:1997:TPC:550448}. This process
algebra has been widely used for the specification and verification of
communicating concurrent control systems. Many of these applications have
been described in~\cite{Roscoe2010} and in the references there.

The complete testing theory presented here applies to CSP processes that are
divergence-free\footnote{The assumption of divergence freedom is usually
applied in black-box testing, since it cannot distinguish between divergence
and deadlock.} and interpreted both in the trace and the failures semantics.
Our results complement work published in~\cite{DBLP:conf/pts/CavalcantiS17}.
There, fault domains are specified as collections of processes refining a
``most general'' fault domain member. With that concept, complete test suites
may be finite or infinite. This gives important insight into the theory of
fault-domain testing for CSP, but we are particularly interested in {\it
finite} suites when it comes to practical application.
While~\cite{DBLP:conf/pts/CavalcantiS17} may require additional criteria to
select tests from still infinite test suites, here, we further restrict fault
domains using a graph representation of processes used in model checking to
obtain test suites that are finite.

Our complementary approach to the definition of CSP fault domains is
presented in this paper. We observe that every finite-state CSP process can
be represented as a finite normalised transition graph, whose edges are
labelled by the events the process engages in, and whose nodes are labelled
by minimal acceptances or, alternatively, maximal
refusals~\cite{Roscoe:1994:CME:197600}. The maximal refusals express the
degree of nondeterminism present in a given process state that is in
one-one-correspondence to a node of the normalised transition graph. Inspired
by the way that fault-domains are specified for finite state machines (FSMs),
we define them as the set of CSP processes whose normalised transition graphs
do not exceed the size of the reference model's graph by more than a given
constant.

The main contribution of this paper is the proof that for fault domains of
the described type, finite, complete test suite generation methods can be
given for testing against trace and failures refinement and equivalence. The
existence of -- possibly infinite -- complete test suites has been
established for process algebras, and for CSP in particular, by several
authors~\cite{Hennessy:1988:ATP:50497,Schneider:1995:OST:203471.203475,DBLP:conf/fm/PeleskaS96,peleska1997a,DBLP:conf/pts/CavalcantiS17}.
To the best of our knowledge, however, this article is the first to present
{\it finite}, complete test suites associated with this class of fault
domains and conformance relations.

Moreover, our result is not a simple transcription of existing knowledge
about finite, complete test suites for finite state machines: the
capabilities of CSP to express nondeterminism in a more fine-grained way than
it is possible for FSMs requires a more complex approach to testing systems
for conformity in the failures-model than it is required for model-based
testing against nondeterministic FSMs, as published, for example,
in~\cite{hierons_testing_2004,DBLP:conf/hase/PetrenkoY14}.

It should be noted that the presented results are of a theoretical
nature:~despite being finite, the resulting test suite size will still be too
large to be applied in practice to real-world problems. We discuss a number
of promising options how the test suite size can be reduced to become
practically applicable.

@@

\subsubsection*{Overview}
In Section~\ref{section:preliminaries},
we present the background material relevant to our work.

@todo

>>>>

In the safety-critical domains, test suites with guaranteed fault coverage
are of particular interest. For black-box testing, guarantees can be given
only if certain hypotheses are satisfied. These hypotheses are usually
specified by a \emph{fault domain}:~a set of models that may or may not
conform to a given reference model. 
The so-called \emph{complete} test strategies guarantee
to accept every  system under test (SUT) conforming to the reference model, while  uncovering every conformance violation, provided that the true SUT behaviour is captured by a member of the
fault domain.

Generation methods for complete test suites have been developed for various
modelling formalisms. In this paper, we use \emph{Communicating Sequential
Processes (CSP)}~\cite{Hoare:1985:CSP:3921,Roscoe2010}. This is a mature
process-algebraic approach that has been shown to be well-suited for the
description of reactive control systems in many publications over almost five
decades. Many of these applications have
been described in~\cite{Roscoe2010} and in the references there.
Industrial success has also been reported, see, for example, \cite{976937,DBLP:conf/prdc/ShiPK99,DBLP:conf/amast/ButhKPS97}.

@@

This paper presents complete black-box test suites for software and systems
modelled using CSP.
They can be generated for divergence-free\footnote{The assumption of divergence freedom is usually
applied in black-box testing, since it cannot distinguish between divergence
and deadlock.} finite-state CSP processes with finite alphabets,   
interpreted both in the trace and the failures semantics.
Our results complement work published in~\cite{DBLP:conf/pts/CavalcantiS17}.
There, fault domains are specified as collections of processes refining a
``most general'' fault domain member. With that concept, complete test suites
may be finite or infinite. This gives important insight into the theory of
fault-domain testing for CSP, but we are particularly interested in {\it
finite} suites when it comes to practical application.
While~\cite{DBLP:conf/pts/CavalcantiS17} may require additional criteria to
select tests from still infinite test suites, here, we further restrict fault
domains using a graph representation of processes (originally elaborated
for the purpose of model checking) to
obtain test suites that are finite.

Our complementary approach to the definition of CSP fault domains is
presented in this paper. We observe that the failures semantics of 
every finite-state CSP process can
be represented as a finite normalised transition graph, whose edges are
labelled by the events the process engages in, and whose nodes are labelled
by minimal acceptances or, alternatively, maximal
refusals~\cite{Roscoe:1994:CME:197600}. The maximal refusals express the
degree of nondeterminism present in a given process state that is in
one-one-correspondence to a node of the normalised transition graph. Inspired
by the way that fault-domains are specified for finite state machines (FSMs),
we define them here as the set of CSP processes whose normalised transition graphs
do not exceed the order (i.e.~the number of nodes) 
of the reference model's graph by more than a given
constant.

The main contribution of this paper consists in the construction of two test suites
verifying the conformance relations \emph{trace refinement} and  \emph{failures refinement}, respectively, for any given non-divergent, finite-state, finite alphabet 
CSP reference process. We prove their completeness with respect to fault domains
of the described type.  The
existence of -- possibly infinite -- complete test suites has been
established for process algebras, and for CSP in particular, by several
authors~\cite{Hennessy:1988:ATP:50497,Schneider:1995:OST:203471.203475,DBLP:conf/fm/PeleskaS96,peleska1997a,DBLP:conf/pts/CavalcantiS17}.
To the best of our knowledge, however, this article is the first to present
{\it finite}, complete test suites associated with this class of fault
domains and conformance relations.

Moreover, our result is not a simple transcription of existing knowledge
about finite, complete test suites for finite state machines: the
capabilities of CSP to express nondeterminism in a more fine-grained way than
it is possible for FSMs requires a more complex approach to testing systems
for conformity in the failures-model than it is required for model-based
testing against nondeterministic FSMs, as published, for example,
in~\cite{hierons_testing_2004,DBLP:conf/hase/PetrenkoY14}. We show that the approach to testing the admissibility of an SUT's nondeterministic behaviour is optimal in the
sense that it cannot be established with fewer test cases.

%It should be noted that the presented results are of a theoretical
%nature:~despite being finite, the resulting test suite size will still be too
%large to be applied in practice to real-world problems. We discuss a number
%of promising options how the test suite size can be reduced to become
%practically applicable.

@@

\subsubsection*{Overview}
In Section~\ref{section:preliminaries},
we present the background material relevant to our work. In Section~\ref{sec:finitecompletefails}, finite complete test suites for verifying failures refinement are presented. Test suites checking trace refinement may be regarded as a simplified version of the former class, they are presented in Section~\ref{sec:finitecomplete}. A sample test suite is presented in Section~\ref{sec:case}. The results obtained are discussed in Section~\ref{sec:conc}.

References to related work are given throughout the paper where appropriate.
